---
title: "Lee-Ann's Part S&DS Final"
author: "Lee-Ann Kao"
date: "2023-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Web Scraping
```{r}
library('rvest')
library(stringr)

#gdp per capita
url <- "https://photius.com/rankings/economy/gdp_per_capita_2007_0.html"
webpage <- read_html(url)
country_gdp <- html_text(html_nodes(webpage, '.desktop~ tr+ tr td:nth-child(3)'))
gdp_per_capita <- as.numeric(gsub("\\$|,", "", html_text(html_nodes(webpage, '.desktop~ tr+ tr td:nth-child(4)'))))
gdp <- data.frame(row.names = country_gdp, gdp_per_capita = gdp_per_capita)

#democracy-dictatorship score
url2 <- "https://en.wikipedia.org/wiki/Democracy-Dictatorship_Index"
webpage2 <- read_html(url2)
country_DD <- html_text(html_nodes(webpage2, '.datasortkey a'))
DD <- html_text(html_nodes(webpage2, 'td:nth-child(2)'))
DD_type <- gsub("\n", "", html_text(html_nodes(webpage2, 'td:nth-child(3)')))
DD_score <- data.frame(row.names = country_DD, democracy_dictatorship_score = DD,    type_of_government = DD_type)

#percent in parliament
url3 <- "http://archive.ipu.org/wmn-e/arc/classif311207.htm"
webpage3 <- read_html(url3)
country_parliament <- gsub(" 1", "", html_text(html_nodes(webpage3, 'tr~ tr+ tr td:nth-child(2)')))
percent_parliament <- as.numeric(gsub("\\%", "", html_text(html_nodes(webpage3, 'tr~ tr+ tr td:nth-child(6)'))))
parliament <- data.frame(row.names = country_parliament, percent_women_in_parliament = percent_parliament)

#male vs. female gross national income per capita (USD PPP)
url4 <- "https://en.wikipedia.org/wiki/List_of_sovereign_states_by_male_to_female_income_ratio"
webpage4 <- read_html(url4)
country_income <- str_trim(gsub("\\*", "", html_text(html_nodes(webpage4, 'td:nth-child(1)'))))
female_income <- as.numeric(gsub(",", "", html_text(html_nodes(webpage4, 'td:nth-child(3)'))))
male_income <- as.numeric(gsub(",", "", html_text(html_nodes(webpage4, 'td:nth-child(2)'))))
ratio <- as.numeric(gsub("\n", "", html_text(html_nodes(webpage4, 'td:nth-child(4)'))))
income <- data.frame(row.names = country_ratio, female_GNI_per_capita = female_income, male_GNI_per_capita = male_income, male_to_female_income_ratio = ratio)

#Human Development Index
url5 <- "https://countryeconomy.com/hdi?year=2007"
webpage5 <- read_html(url5)
country_hdi <- str_trim(gsub("\\[|\\+|\\]", "", html_text(html_nodes(webpage5, 'td:nth-child(1)'))))
hdi_score <- as.numeric(html_text(html_nodes(webpage5, '.numero:nth-child(2)')))
hdi_tiers <- cut(hdi_scores, breaks = c(0, 0.55, 0.7, 0.8, 1), labels = c("low", "medium", "high", "very high"))
hdi <- data.frame(row.names = country_hdi, hdi_score = hdi_score, hdi_tiers = hdi_tiers)

#merge
merge1 <- merge(gdp, DD_score, by = 'row.names', all = TRUE)
rownames(merge1) <- merge1$Row.names
merge2 <- merge(parliament, income, by = 'row.names', all = TRUE)
rownames(merge2) <- merge2$Row.names
merge3 <- merge(merge2, hdi, by = 'row.names', all = TRUE)
rownames(merge3) <- merge3$Row.names
final_merge <- merge(merge1, merge3, by = 'row.names', all = TRUE)
rownames(final_merge) <- final_merge$Row.names
final_merge <- final_merge[, c(3:5, 8:13)]
final_merge <- final_merge[complete.cases(final_merge), ]
```
*I scraped data on GDP per capita, Democracy-Dictatorship classifications, government type, percent of women in lower/single house parliament, male vs. female gross national income per capita (USD PPP), and Human Development Index for countries in 2007 across various websites with the help of the SelectorGadget Chrome extension. I then cleaned the data as necessary using gsub for commas, trailing spaces, and other characters, made data frames of each variable, merged all data frames by country name, and took only complete cases.*


#One-Way ANOVA
```{r, show_col_types = FALSE}
gender_info <- readr::read_csv("https://raw.github.com/ava-chae/genderinfo/main/finalmerge.csv")
gender_info <- as.data.frame(gender_info[-46, ])
gender_info$womens_share_of_tertiary_enrolment_percent <- as.numeric(gender_info$womens_share_of_tertiary_enrolment_percent)
library(car)
library(leaps)
```
```{r}
#boxplot
par(mar = c(10, 4, 5, 2), cex = .8)
boxplot(womens_share_of_tertiary_enrolment_percent ~ type_of_government, data = gender_info, col = 'yellow', main = "Women's Share of Tertiary Enrollment by Government Type", ylab = "Women's Share of Tertiary Enrollment", xlab = "", cex.axis = 0.7, las = 2)
means <- tapply(gender_info$womens_share_of_tertiary_enrolment_percent, gender_info$type_of_government, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x = c(1:6), y = means + 4, labels = round(means,2))

#ratio of standard deviations
(sds <- tapply(gender_info$womens_share_of_tertiary_enrolment_percent, gender_info$type_of_government, sd))
round(max(sds)/min(sds), 1)

#Fit ANOVA model - use the aov function
aov1 <- aov(gender_info$womens_share_of_tertiary_enrolment_percent ~ gender_info$type_of_government)
summary(aov1)

#same thing using lm() - basically regression with indicator variables.
mod1 <- lm(gender_info$womens_share_of_tertiary_enrolment_percent ~ gender_info$type_of_government -1)
summary(mod1)
CIs <- confint(mod1)
round(CIs, 2)
coefs <- coef(mod1)

#plot
library(plotrix)
par(cex.main = 0.7, cex.lab = 0.8, mar = c(4, 12, 5, 2))
plotCI(coefs, 1:(length(coefs)), ui = CIs[,2], li = CIs[,1], axes = FALSE, err = "x", ylab = "", xlab = "Mean (and 95% CI)", main = "Mean and CI's for Women's Share of Tertiary Enrollment by Government Type", lwd = 2, col = "blue")
axis(side = 1)
axis(side = 2, at = 1:(length(coefs)), label = levels(as.factor(gender_info$type_of_government)), las = 2)

#Tukey
TukeyHSD(aov1)
par(cex.main = 0.75, cex.axis = 0.5)
plot(TukeyHSD(aov1), las = 1)

#check assumptions
par(cex.main = 0.75, cex.lab = 1, mar = c(4, 5, 5, 2))
myResPlots2(aov1, label = "Women's Share of Tertiary Enrollment by Government Type")
bartlett.test(gender_info$womens_share_of_tertiary_enrolment_percent, gender_info$type_of_government)
leveneTest(gender_info$womens_share_of_tertiary_enrolment_percent, gender_info$type_of_government)
```
*From the boxplot it appears that there is a relationship between women's share of tertiary education and government type because the five-point-summary of fatality varies greatly by government type (in particular, civilian and military dictatorships have a much smaller women's share of tertiary enrollment). It is reasonable to assume that the variances are the same across groups because the ratio of largest to smallest sample standard deviation is 1.7, which is less than 2. The mean share of tertiary enrollment is statistically significantly different between government types because the p-value of 0.000147 is less than alpha at 0.05, so we can reject the null hypothesis that there is no difference in mean women's share of tertiary education by government type. Except for in comparison to Royal Dictatorships, any pair of government type with Military Dictatorships had a statistically significant difference in mean women's share of tertiary enrollment, as the p-values were all less than alpha at 0.05. Finally, the residuals appear to be normally distributed as most fall within the 95% confidence interval of the normal quantile plot. Furthermore, the residuals all fall within the 95% confidence interval in the fits vs. residuals plot, and our tests of equal variance do not show a statistically significant difference in standard deviations across groups, as the p-values are greater than alpha at 0.05. Therefore, we meet ANOVA's assumptions that residuals are normally distributed within each group and that variability within groups is relatively constant across groups.*
