---
title: 'Gender Equality in 2007: An Analysis Across Countries and Governments'
author: "Long Live the Queens"
date: "2023-05-06"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(car)
library(leaps)
library('rvest')
library(stringr)
library(dplyr)
library(plotrix)
source("http://www.reuningscherer.net/s&ds230/Rfuncs/regJDRS.txt")
```

```{r, include = FALSE}
permCor <- function(x, y, n_samp = 10000, plotit = T){
   corResults <- rep(NA, n_samp)
   for (i in 1:n_samp){
      corResults[i] <- cor(x, sample(y))
   }
   
   pval <- mean(abs(corResults) >= abs(cor(x, y)))
   
   if (plotit == T){
      #Make histogram of permuted correlations
      hist(corResults, col = "yellow", main = "", xlab = "Correlations", breaks = 50,
           xlim = range(corResults,cor(x,y)))
      mtext("Permuted Sample Correlations", cex = 1.2, line = 1)
      mtext(paste0("Permuted P-value = ", round(pval, 4),", Calculated P-value = ", round(cor.test(x, y)$p.value, 4)), cex = 0.8, line = 0)
      abline(v = cor(x, y), col = "blue", lwd = 3)
      text(cor(x,y)*1.05, 0, paste("Actual Correlation =", round(cor(x, y), 2)), srt = 90, adj = 0)
   }
   if (plotit == F){
      return(round(pval, 5))
   }  
}
```

#### Introduction
*Insert text (background, motivation - not more than a short paragraph)*

#### Web Scraping
```{r}
#gdp per capita
url <- "https://photius.com/rankings/economy/gdp_per_capita_2007_0.html"
webpage <- read_html(url)
country_gdp <- html_text(html_nodes(webpage, '.desktop~ tr+ tr td:nth-child(3)'))
gdp_per_capita <- as.numeric(gsub("\\$|,", "", html_text(html_nodes(webpage, '.desktop~ tr+ tr td:nth-child(4)'))))
gdp <- data.frame(row.names = country_gdp, gdp_per_capita = gdp_per_capita)
#democracy-dictatorship score
url2 <- "https://en.wikipedia.org/wiki/Democracy-Dictatorship_Index"
webpage2 <- read_html(url2)
country_DD <- html_text(html_nodes(webpage2, '.datasortkey a'))
DD <- html_text(html_nodes(webpage2, 'td:nth-child(2)'))
DD_type <- gsub("\n", "", html_text(html_nodes(webpage2, 'td:nth-child(3)')))
DD_score <- data.frame(row.names = country_DD, democracy_dictatorship_score = DD,    type_of_government = DD_type)
#percent in parliament
url3 <- "http://archive.ipu.org/wmn-e/arc/classif311207.htm"
webpage3 <- read_html(url3)
country_parliament <- gsub(" 1", "", html_text(html_nodes(webpage3, 'tr~ tr+ tr td:nth-child(2)')))
percent_parliament <- as.numeric(gsub("\\%", "", html_text(html_nodes(webpage3, 'tr~ tr+ tr td:nth-child(6)'))))
parliament <- data.frame(row.names = country_parliament, percent_women_in_parliament = percent_parliament)
#male vs. female gross national income per capita (USD PPP)
url4 <- "https://en.wikipedia.org/wiki/List_of_sovereign_states_by_male_to_female_income_ratio"
webpage4 <- read_html(url4)
country_income <- str_trim(gsub("\\*", "", html_text(html_nodes(webpage4, 'td:nth-child(1)'))))
female_income <- as.numeric(gsub(",", "", html_text(html_nodes(webpage4, 'td:nth-child(3)'))))
male_income <- as.numeric(gsub(",", "", html_text(html_nodes(webpage4, 'td:nth-child(2)'))))
ratio <- as.numeric(gsub("\n", "", html_text(html_nodes(webpage4, 'td:nth-child(4)'))))
income <- data.frame(row.names = country_income, female_GNI_per_capita = female_income, male_GNI_per_capita = male_income, male_to_female_income_ratio = ratio)
#merge
merge1 <- merge(gdp, DD_score, by = 'row.names', all = TRUE)
rownames(merge1) <- merge1$Row.names
merge2 <- merge(parliament, income, by = 'row.names', all = TRUE)
rownames(merge2) <- merge2$Row.names
final_merge <- merge(merge1, merge2, by = 'row.names', all = TRUE)
rownames(final_merge) <- final_merge$Row.names
final_merge <- final_merge[, c(3:5, 7:10)]
final_merge <- final_merge[complete.cases(final_merge), ]
```
*I scraped data on GDP per capita, Democracy-Dictatorship classifications, government type, percent of women in lower/single house parliament, and male vs. female gross national income per capita (USD PPP) for countries in 2007 across various websites with the help of the SelectorGadget Chrome extension. I then cleaned the data as necessary using gsub for commas, trailing spaces, and other characters, made data frames of each variable, merged all data frames by country name, and took only complete cases.*

#### Data Retrieving & Cleaning

```{r, warning = F, message = F}
# read csv file from github to gendergf
genderdf <- readr::read_csv("https://raw.github.com/ava-chae/genderinfo/main/genderinfocondensed.csv")
# cut dataframe to include only 6 variables
gendertrial <- subset(genderdf, genderdf$category %in% c("total_fertility_rate", "womens_share_of_labour_force", "womens_share_of_tertiary_enrolment", "labour_force_participation_rate", "singulate_mean_age_at_marriage", "contraceptive_prevalence_rate_any_method"))
# only keep first four characters in year
gendertrial$year <- substr(gendertrial$year, 0, 4)
```

```{r, warning = F, message = F}
gendertrial <- gendertrial[,-c(4, 6)]
gendertrial <- gendertrial[complete.cases(gendertrial), ]
gendertrial <- gendertrial[gendertrial$value != "…", ]
gendertrial <- gendertrial[order(gendertrial[,3], decreasing = TRUE), ]
gendertrial <- gendertrial[!duplicated(gendertrial[c("country_or_area", "category", "subgroup")]),]
```

```{r, warning = F, message = F}
# make total fertility rate subset
tfr_df <- subset(gendertrial, gendertrial$category == 'total_fertility_rate', select = c('country_or_area', 'value'))
colnames(tfr_df)[which(names(tfr_df) == "value")] <- "fertility_rate"
# make women's share of labour force subset
wslf_df <- subset(gendertrial, gendertrial$category == 'womens_share_of_labour_force', select = c('country_or_area', 'value'))
colnames(wslf_df)[which(names(wslf_df) == "value")] <- "w_labor_force"
# make women's share of tertiary enrollment subset
wste_df <- subset(gendertrial, gendertrial$category == 'womens_share_of_tertiary_enrolment', select = c('country_or_area', 'value'))
colnames(wste_df)[which(names(wste_df) == "value")] <- "w_tert_enroll"
# make labor force participation rate subset split by gender
lfp_f_df <- subset(gendertrial, gendertrial$category == 'labour_force_participation_rate' & gendertrial$subgroup == 'Female 15+ yr', select = c('country_or_area', 'value'))
colnames(lfp_f_df)[which(names(lfp_f_df) == "value")] <- "lfp_rate_f"
lfp_m_df <- subset(gendertrial, gendertrial$category == 'labour_force_participation_rate' & gendertrial$subgroup == 'Male 15+ yr', select = c('country_or_area', 'value'))
colnames(lfp_m_df)[which(names(lfp_m_df) == "value")] <- "lfp_rate_m"
# make singulate mean age at marriage subset split by gender
age_mar_f_df <- subset(gendertrial, gendertrial$category == 'singulate_mean_age_at_marriage' & gendertrial$subgroup == 'Female', select = c('country_or_area', 'value'))
colnames(age_mar_f_df)[which(names(age_mar_f_df) == "value")] <- "age_marriage_f"
age_mar_m_df <- subset(gendertrial, gendertrial$category == 'singulate_mean_age_at_marriage' & gendertrial$subgroup == 'Male', select = c('country_or_area', 'value'))
colnames(age_mar_m_df)[which(names(age_mar_m_df) == "value")] <- "age_marriage_m"
# make contraceptive prevalence rate any method subset
contprev_df <- subset(gendertrial, gendertrial$category == 'contraceptive_prevalence_rate_any_method', select = c('country_or_area', 'value'))
colnames(contprev_df)[which(names(contprev_df) == "value")] <- "cont_prev_rate"
```

```{r, warning = F, message = F}
# all subsets merged 
allmerge0 = merge(lfp_f_df, merge(lfp_m_df, merge(age_mar_f_df, merge(age_mar_m_df, merge(contprev_df, merge(wste_df, merge(wslf_df, tfr_df, by = c("country_or_area"), all = T), by = c("country_or_area"), all = T), by = c("country_or_area"), all = T), by = c("country_or_area"), all = T), by = c("country_or_area"), all = T), by = c("country_or_area"), all = T), by = c("country_or_area"), all = T)
completemerge0 <- allmerge0[complete.cases(allmerge0),]
```

```{r, warning = F, message = F}
# read csv file from github to scraped data
scrapeddf <- readr::read_csv("https://raw.github.com/ava-chae/genderinfo/main/*UPDATED_scrapedmerge.csv")
colnames(scrapeddf)[1] = "country_or_area"
# merge scraped data with gender data
finalmerge <- merge(completemerge0, scrapeddf, by = c("country_or_area"), all = T)
gender_info <- finalmerge[complete.cases(finalmerge),]
# make as numeric
gender_info <- gender_info %>% mutate_at(c('lfp_rate_f', 'lfp_rate_m', 'age_marriage_m', 'age_marriage_f', 'cont_prev_rate', 'w_tert_enroll', 'w_labor_force', 'fertility_rate'), as.numeric)
```
*The final cleaned and merged dataset used 14 continuous variables and 3 categorical variables. Gender-specific data was collected from Kaggle (https://www.kaggle.com/datasets/unitednations/gender-info-2007) and additional data was scraped from various websites.*

*The original Kaggle gender information data was first condensed by removing irrelevant columns to be able to upload to GitHub. The condensed datafile is loaded then cut to include only 6 variables that we wanted to analyze. The values of year sometimes had multiple years, so for simplicity, we only kept the first year in each row.  Irrelevant columns were removed, and only complete rows were kept. Next, because some countries had several years worth of data for certain variables, we only kept the most recent data values for duplicate rows. The major cleaning aspect was splitting and merging the dataset. The dataset did not have each variable as a separate column and instead just named all the variables in one column, we had to separate the rows to make analysis much easier. Each variable was split and merged together to form a dataframe in which each variable was given a separate column. Only complete cases were kept in the final data. Finally, the additional web scraped data was pulled from GitHub and merged with the cleaned Kaggle data to form the finalmerge.csv file for analysis. The main issues we encountered was finding that data for most variables only existed for certain countries as well as separating variables into different columns without taking too many lines of code.*

* Labor force participation rate (lfp_rate_m & f): continuous as a percent
* Mean age at marriage (age_marriage_m & f): continuous in years
* Contraceptive prevalence rate (cont_prev_rate): continuous as percent
* Women's share of tertiary enrollment (w_tert_enroll): continuous as percent
* Women's share of labor force (w_labor_force): continuous as percent
* Total fertility rate per woman (fertility_rate): continuous as number of average births per woman
* GDP per capita (in 2007): continuous average GDP per person
* Percent women in parliament: continuous as percent of 100
* Male to female income ratio: continuous average male to female income ratio
* GNI per capita (male and female): continuous average GNI per person
* Type of government: categorical classification of national government type

#### Correlation Test 
```{r}
class(gender_info$cont_prev_rate)
(cor1test <- cor.test(gender_info$cont_prev_rate, gender_info$fertility_rate))
plot(gender_info$cont_prev_rate, gender_info$fertility_rate, pch = 19, col = "skyblue", xlab = "Contraceptive Prevalence Rate", ylab = "Total Fertility Rate (per woman)")
mtext("Contraceptive Prevalence Rate vs Total Fertility Rate", cex = 1.2, line = 1)
```
*The p-value of < 2.2e-16 is extremely small (certainly less than 0.05), so we conclude that the correlation is statistically significantly different from zero. The confidence interval for the true correlation is (-0.82, -0.92) which is well away from zero.*

#### T-test & T-test Bootstrap
(Show ability to create bootstrap confidence interval for either a t-test or a correlation. For full marks, should include histogram with both regular and bootstrapped CI shown, as well as a short sentence description of results. Plots should be clearly labeled, well formatted, and display an aesthetic sense.)
```{r}
t.test(gender_info$percent_women_in_parliament, mu = 0.5)
boxplot(gender_info$percent_women_in_parliament, horizontal = T, col = "royalblue", xlab = "Percentage", main = "Percentage of Women in Parliament", lwd = 2)
(parm_ci <- t.test(gender_info$percent_women_in_parliament, mu = 0.5)$conf.int) 
N <- nrow(gender_info)
n_samp <- 10000
WomenParli <- rep(NA, n_samp)
for (i in 1:n_samp){
sample_wparli <- sample(gender_info$percent_women_in_parliament, length(gender_info$percent_women_in_parliamen), replace = TRUE)  
WomenParli[i] <- mean(sample_wparli)
}
(boot_ci <- quantile(WomenParli, c(0.025, 0.975)))
hist(gender_info$percent_women_in_parliament, main = "Percent of Women in Parliament", xlab = "Percent", col = "royalblue", cex.main = 0.8, breaks = 50)
abline(v = boot_ci,col = "palevioletred",lwd = 3)
abline(v = parm_ci, col = "slategray2", lwd = 3, lty = 2)
legend("topleft", c("Parametric CI", "Bootstrapped CI"), lwd = 3, col = c("brown2","purple2"), lty = c(2, 1))
hist(WomenParli, main = "Mean Percent Return Percent of Women in Parliament", xlab = "Percent", col = "royalblue", cex.main = 0.8, breaks = 50)
abline(v = boot_ci,col = "palevioletred",lwd = 3)
abline(v = parm_ci, col = "slategray2", lwd = 3, lty = 2)
legend("topleft", c("Parametric CI", "Bootstrapped CI"), lwd = 3, col = c("slategray2","palevioletred"), lty = c(2, 1))
```
*Given that the p-value is less than 2.2e-16, we can reject the null hypothesis. We reject the null hypothesis that the true mean of women’s share of tertiary education is equal to 50%.*
*The boot CI is a bit narrower than the parametric CI; however, the difference is quite small.*

#### Permutation Test 
(Include at least one. For full marks, should include histogram with results, line for actual statistic value, p-value of permutation, and short sentence description of results)
```{r}
permCor(gender_info$lfp_rate_m, gender_info$lfp_rate_f)
permCor(gender_info$lfp_rate_m,gender_info$lfp_rate_f, plotit = F)
par.cex = 0.6
permCor(gender_info$male_GNI_per_capita, gender_info$female_GNI_per_capita) 
permCor(gender_info$male_GNI_per_capita, gender_info$female_GNI_per_capita, plotit = F)
```
*The permuted sample correlations for labor force participation rate for males and females has a p-value of 0.0176. This suggests that we reject the null hypothesis. At a significance level of 0.05, we conclude that the labor force participation rate for men and women are different between gender groups.The permuted sample correlations for GNI per capita between men and women has a p-value of 0. We reject the null hypothesis as the observed data is extremely unlikely to have occurred by chance alone. Therefore, the men and women GNI per capita difference is highly based off the difference in gender.* 

#### Multiple Regression
(Few sentence intro to plan for multiple regression, variables used, process, etc. Use either backwards stepwise regression or some form of best subsets regression. Should include residual plots. A GLM with a mix of continuous and categorical predictors is fine here.)
*Insert plan*
(Normal Quantile Plot of residuals, plot of fits vs. residuals, discussion. Plots should be clearly labeled, well formatted, and display an aesthetic sense.)
```{r}
hist(gender_info$gdp_per_capita, col = "palevioletred", main = "GDP Per Capita")
qqPlot(gender_info$gdp_per_capita, main = "QQ Plot GDP Per Capita", pch = 19)

#lambda = 0 means log transformation
gender_info3 <- na.omit(gender_info[ , c("gdp_per_capita", "lfp_rate_f", "lfp_rate_m", "age_marriage_f", "age_marriage_m", "cont_prev_rate", "w_tert_enroll", "fertility_rate", "percent_women_in_parliament", "male_to_female_income_ratio")])
linear_m <- lm(gdp_per_capita ~ ., data = gender_info3)
trans1 <- boxCox(linear_m)
trans1$x[which.max(trans1$y)]

#Make variable on log scale for more normal data
gender_info$logGDP <- log(gender_info$gdp_per_capita)
hist(gender_info$logGDP, col = "slateblue", main = "Log GDP Per Capita")
qqPlot(gender_info$logGDP, main = "QQ Plot Log GDP Per Capita", pch = 19)

gender_info2 <- na.omit(gender_info[ , c("logGDP", "lfp_rate_f", "lfp_rate_m", "age_marriage_f", "age_marriage_m", "cont_prev_rate", "w_tert_enroll", "fertility_rate", "percent_women_in_parliament", "male_to_female_income_ratio")])
rownames(gender_info2) <- gender_info[, 1]

suppressWarnings(pairsJDRS(gender_info2))
#issues of multi-collinearity as the predictors are correlated with eachother
#variables will be 'competing' with eachother to explain the variability in log of GDP


#Backwards Stepwise Regression
lm1 <- lm(logGDP ~ ., data = gender_info2)
summary(lm1)
lm2 <- lm(logGDP ~ lfp_rate_f +  age_marriage_f + fertility_rate , data = gender_info2)
summary (lm2)

#NQ Plot and Fits vs. Residual Plot
myResPlots2(lm2, "Model for log GDP per capita")
```
*Insert discussion of criteria used to choose final model-- Results of final model displayed. Discussion of R-squared, parameter direction and significance. Discussion of criteria used to choose final model*
*Insert discussion of results (Transformations, heteroskedasticity, completeness, etc.)*

#### ANOVA
For ANOVA, should have model, interaction plot if two way, discussion of results, Tukey comparisons, residual plots. 
```{r}
#boxplot
par(mar = c(10, 4, 5, 2), cex = .8)
boxplot(w_tert_enroll ~ type_of_government, data = gender_info, col = 'yellow', main = "Women's Share of Tertiary Enrollment by Government Type", ylab = "Women's Share of Tertiary Enrollment", xlab = "", cex.axis = 0.7, las = 2)
means <- tapply(gender_info$w_tert_enroll, gender_info$type_of_government, mean)
points(means, col = "red", pch = 19, cex = 1.2)
text(x = c(1:6), y = means + 4, labels = round(means,2))
#ratio of standard deviations
(sds <- tapply(gender_info$w_tert_enroll, gender_info$type_of_government, sd))
round(max(sds)/min(sds), 1)
#Fit ANOVA model - use the aov function
aov1 <- aov(gender_info$w_tert_enroll ~ gender_info$type_of_government)
summary(aov1)
#same thing using lm() - basically regression with indicator variables.
mod1 <- lm(gender_info$w_tert_enroll ~ gender_info$type_of_government -1)
summary(mod1)
CIs <- confint(mod1)
round(CIs, 2)
coefs <- coef(mod1)
#plot
par(cex.main = 0.7, cex.lab = 0.8, mar = c(4, 12, 5, 2))
plotCI(coefs, 1:(length(coefs)), ui = CIs[,2], li = CIs[,1], axes = FALSE, err = "x", ylab = "", xlab = "Mean (and 95% CI)", main = "Mean and CI's for Women's Share of Tertiary Enrollment by Government Type", lwd = 2, col = "blue")
axis(side = 1)
axis(side = 2, at = 1:(length(coefs)), label = levels(as.factor(gender_info$type_of_government)), las = 2)
#Tukey
TukeyHSD(aov1)
par(cex.main = 0.75, cex.axis = 0.5)
plot(TukeyHSD(aov1), las = 1)
#check assumptions
par(cex.main = 0.75, cex.lab = 1, mar = c(4, 5, 5, 2))
myResPlots2(aov1, label = "Women's Share of Tertiary Enrollment by Government Type")
bartlett.test(gender_info$w_tert_enroll, gender_info$type_of_government)
leveneTest(gender_info$w_tert_enroll, gender_info$type_of_government)
```
*Insert discussion of results*

#### Conclusion
(Quality of presentation, misc additional work, brevity, understanding of material.)
*Insert Conclusion - short paragraph*