---
title: "S&DS Final Multiple Regression"
author: "Jennifer Centa"
date: "2023-05-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(car)
library(leaps)
source("http://www.reuningscherer.net/s&ds230/Rfuncs/regJDRS.txt")
```

## R Markdown

```{r}
(gender_info <- readr::read_csv("https://raw.github.com/ava-chae/genderinfo/main/finalmerge.csv", show_col_types = FALSE))
gender_info <- as.data.frame(gender_info[-46, ])
```

* bc of collinearity, we used backwards stepwise instead of forwards

## Description of Plan (backwards stepwise regression or best subsets regression & include residual plots)

```{r}
hist(gender_info$gdp_per_capita, col = "palevioletred", main = "GDP Per Capita")
qqPlot(gender_info$gdp_per_capita, main = "QQ Plot GDP Per Capita", pch = 19)

#lambda = 0 means log transformation
gender_info3 <- na.omit(gender_info[ , c("gdp_per_capita", "labour_force_participation_rate_female_percent", "labour_force_participation_rate_male_percent", "mean_age_at_marriage_female", "mean_age_at_marriage_male", "contraceptive_prevalence_rate_any_method_percent", "womens_share_of_tertiary_enrolment_percent", "total_fertility_rate_per_woman", "percent_women_in_parliament", "male_to_female_income_ratio")])
linear_m <- lm(gdp_per_capita ~ ., data = gender_info3)
trans1 <- boxCox(linear_m)
trans1$x[which.max(trans1$y)]

#Make variable on log scale for more normal data
gender_info$logGDP <- log(gender_info$gdp_per_capita)
hist(gender_info$logGDP, col = "slateblue", main = "Log GDP Per Capita")
qqPlot(gender_info$logGDP, main = "QQ Plot Log GDP Per Capita", pch = 19)

gender_info2 <- na.omit(gender_info[ , c("logGDP", "labour_force_participation_rate_female_percent", "labour_force_participation_rate_male_percent", "mean_age_at_marriage_female", "mean_age_at_marriage_male", "contraceptive_prevalence_rate_any_method_percent", "womens_share_of_tertiary_enrolment_percent", "total_fertility_rate_per_woman", "percent_women_in_parliament", "male_to_female_income_ratio")])
gender_info2$womens_share_of_tertiary_enrolment_percent <- as.numeric(gender_info2$womens_share_of_tertiary_enrolment_percent)
rownames(gender_info2) <- gender_info[, 1]
```

```{r}
suppressWarnings(pairsJDRS(gender_info2))
```
* issues of multi-collinearity as the predictors are correlated with eachother
* variables will be 'competing' with eachother to explain the variability in log of GDP

## Backwards Stepwise Regression

```{r}
#complete backwards stepwise regression
lm1 <- lm(logGDP ~ ., data = gender_info2)
summary(lm1)

lm2 <- lm(logGDP ~ labour_force_participation_rate_female_percent +  mean_age_at_marriage_female + total_fertility_rate_per_woman , data = gender_info2)
summary (lm2)

#mean_age_at_marriage_male + contraceptive_prevalence_rate_any_method_percent + labour_force_participation_rate_male_percent + percent_women_in_parliament + male_to_female_income_ratio + womens_share_of_tertiary_enrolment_percent 
```

## Multiple Regression Residual Plots (NQ Plot, plot of fits vs. residuals)

```{r}
myResPlots2(lm6, "Model for log GDP per capita")
```
Overall, things look good:

*  Residuals seem approximately normally distributed
*  No evidence of heteroskedasticity
*  ONE OUTLIER
*  No extremely influential points

THE END
